Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-02-07 LaTeX/main.tex
moved appendices out of main text but can crossref

@@

\newpage

\setcounter{equation}{0}
\renewcommand{\theequation}{S1.\arabic{equation}}

\section*{S1 Appendix}

\subsection*{Recursion equations}

In each generation we census the genotype frequencies in male and female gametes/gametophytes (hereafter, gametes) between meiosis (and any meiotic drive) and gametic competition. 
At this stage we denote the frequencies of X- and Y-bearing gametes from males and females $x_{i}^{\Hermaphrodite}$ and $y_{i}^{\Hermaphrodite}$.
The superscript $\Hermaphrodite \in \{\male,\female\}$ specifies the sex of the diploid that the gamete came from. 
The subscript $i\in\{1,2,3,4\}$ specifies the genotype at the selected locus $\mathbf{A}$ and at the novel sex-determining locus $\mathbf{M}$, where $1=AM$, $2=aM$, $3=Am$, and $4=am$. 
The gamete frequencies from each sex sum to one, $\sum_{i}x_{i}^{\Hermaphrodite}+y_{i}^{\Hermaphrodite}=1$.

Competition then occurs among gametes of the same sex (e.g., among eggs and among sperm separately) according to the genotype at the $\mathbf{A}$ locus ($w_{1}^\Hermaphrodite=w_{3}^\Hermaphrodite=w_{A}^\Hermaphrodite$, $w_{2}^\Hermaphrodite=w_{4}^\Hermaphrodite=w_{a}^\Hermaphrodite$, see Table \ref{tab:fitnesstable}).
The genotype frequencies after gametic competition are $x_{i}^{\Hermaphrodite,s}= w_{i}^{\Hermaphrodite}x_{i}^{\Hermaphrodite}/\bar{w}_{H}^{\Hermaphrodite}$ and $y_{i}^{\Hermaphrodite,s}= w_{i}^{\Hermaphrodite}y_{i}^{\Hermaphrodite}/\bar{w}_{H}^{\Hermaphrodite}$, where $\bar{w}_{H}^{\Hermaphrodite}=\sum_{i} w_{i}^{\Hermaphrodite}x_{i}^{\Hermaphrodite}+w_{i}^{\Hermaphrodite}y_{i}^{\Hermaphrodite}$ is the mean fitness of male ($\Hermaphrodite=\male$) or female ($\Hermaphrodite=\female$) gametes.

Random mating then occurs between gametes to produce diploid zygotes.
%To shorten notation we now use index $i$ (and $j$) to denote the alleles at both the $\mathbf{A}$ and $\mathbf{M}$ loci and label $MA=1$, $Ma=2$, $mA=3$, and $ma=4$, such that $i,j\in\{1,2,3,4\}$.
The frequencies of XX zygotes are then denoted as $xx_{ij}$, XY zygotes as $xy_{ij}$, and YY zygotes as $yy_{ij}$, where $\mathbf{A}$ and $\mathbf{M}$ locus genotypes are given by $i,j\in\{1,2,3,4\}$, as above. 
In XY zygotes, the haplotype inherited from an X-bearing gamete is given by $i$ and the haplotype from a Y-bearing gamete is given by $j$. 
In XX and YY zygotes, individuals with diploid genotype $ij$ are equivalent to those with diploid genotype $ji$; for simplicity, we use $xx_{ij}$ and $yy_{ij}$ with $i\neq j$ to denote the average of these frequencies, $xx_{ij}=(x_{i}^{\female,s}x_{j}^{\male,s}+x_{j}^{\female,s}x_{i}^{\male,s})/2$ and $yy_{ij}=(y_{i}^{\female,s}y_{j}^{\male,s}+y_{j}^{\female,s}y_{i}^{\male,s})/2$.

Denoting the $\mathbf{M}$ locus genotype by $b \in \{MM, Mm, mm\}$ and the $\mathbf{X}$ locus genotype by $c \in \{XX, XY, YY\}$, zygotes develop as females with probability $k_{bc}$. 
Therefore, the frequencies of XX females are given by $xx_{ij}^{\female}=k_{bc}xx_{ij}$, XY females are given by $xy_{ij}^{\female}=k_{bc}xy_{ij}$, and YY females are given by $yy_{ij}^{\female}=k_{bc}yy_{ij}$. 
Similarly, XX male frequencies are $xx_{ij}^{\male}=(1-k_{bc})xx_{ij}$, XY male frequencies are $xy_{ij}^{\male}=(1-k_{bc})xy_{ij}$, and YY males frequencies are $yy_{ij}^{\male}=(1-k_{bc})yy_{ij}$.
This notation allows both the ancestral and novel sex-determining regions to determine zygotic sex according to an XY system, a ZW system, or an environmental sex-determining system. 
In addition, we can consider any epistatic dominance relationship between the two sex-determining loci. 
Here, we assume that the ancestral sex-determining system ($\mathbf{X}$ locus) is XY ($k_{MMXX}=1$ and $k_{MMXY}=k_{MMYY}=0$) or ZW ($k_{MMZZ}=0$ and $k_{MMZW}=k_{MMWW}=1$) and epistatically recessive to a dominant novel sex-determining locus, $\mathbf{M}$ ($k_{Mmc}=k_{mmc}=k$).

Selection among diploids then occurs according to the diploid genotype at the $\mathbf{A}$ locus, $l \in \{AA, Aa, aa\}$, for an individual of type $ij$ (see Table \ref{tab:fitnesstable}). 
The diploid frequencies after selection in sex $\Hermaphrodite$ are given by $xx_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} xx_{ij}/\bar{w}^{\Hermaphrodite}_{D}$, $xy_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} xy_{ij}/\bar{w}^{\Hermaphrodite}_{D}$, and $yy_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} yy_{ij}/\bar{w}^{\Hermaphrodite}_{D}$, where $\bar{w}^{\Hermaphrodite}_{D}= \sum_{i=1}^{4}\sum_{j=1}^{4}w_{l}^{\Hermaphrodite}xx_{ij}+w_{l}^{\Hermaphrodite}xy_{ij}+w_{l}^{\Hermaphrodite}yy_{ij}$ is the mean fitness of diploids of sex $\Hermaphrodite$.

Finally, these diploids undergo meiosis to produce the next generation of gametes. 
Recombination and sex-specific meiotic drive occur during meiosis.
%We can also assume that recombination is sex specific and/or affected by the M locus - but generally we don't so I just describe $R=R_{f}=R, r_{MM,d}=r_{Mm,d}=r_{mm,d}=r, \chi_{m}=\chi_{f}=\chi$.
Here, we allow any relative locations for the $\mathbf{X}$, $\mathbf{A}$, and $\mathbf{M}$ loci by using three parameters to describe the recombination rates between them. 
$R$ is the recombination rate between the $\mathbf{A}$ and $\mathbf{M}$ loci, $\rho$ is the recombination rate between the $\mathbf{M}$ and $\mathbf{X}$ loci, and $r$ is the recombination rate between the $\mathbf{A}$ and $\mathbf{X}$ loci (Fig \ref{fig:model_outline}). 
\nameref{tab:chisubstitutions} shows replacements that can be made for each possible ordering of the loci assuming that there is no cross-over interference.
During meiosis in sex $\Hermaphrodite$, meiotic drive occurs such that, in $Aa$ heterozygotes, a fraction $\alpha^{\Hermaphrodite}$ of gametes produced carry the $A$ allele and $(1-\alpha^\Hermaphrodite)$ carry the $a$ allele.

Among gametes from sex $\Hermaphrodite$, the frequencies of haplotypes (before gametic competition) in the next generation are given by

\begingroup
\allowdisplaybreaks
\begin{subequations}
\begin{align}

@@

\begin{split}
x_{1}^{{\Hermaphrodite}'}=&xx_{11}^{\Hermaphrodite,s}+xx_{13}^{\Hermaphrodite,s}/2+(xx_{12}^{\Hermaphrodite,s}+xx_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{14}^{\Hermaphrodite,s}-xx_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&+(xy_{11}^{\Hermaphrodite,s}+xy_{13}^{\Hermaphrodite,s})/2+(xy_{12}^{\Hermaphrodite,s}+xy_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{12}^{\Hermaphrodite,s}-xy_{21}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \rho(xy_{13}^{\Hermaphrodite,s}-xy_{31}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho)xy_{14}^{\Hermaphrodite,s} +(R+\rho-r)xy_{41}^{\Hermaphrodite,s}\\
&+(R+r-\rho) xy_{23}^{\Hermaphrodite,s}+(R+\rho-r)xy_{32}^{\Hermaphrodite,s}\big{]}\alpha^\Hermaphrodite/2
\end{split}
\\

@@

\begin{split}
x_{2}^{{\Hermaphrodite}'}=&xx_{22}^{\Hermaphrodite,s}+xx_{24}^{\Hermaphrodite,s}/2+(xx_{12}^{\Hermaphrodite,s}+xx_{23}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{23}^{\Hermaphrodite,s}-xx_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{22}^{\Hermaphrodite,s}+xy_{24}^{\Hermaphrodite,s})/2+(xy_{21}^{\Hermaphrodite,s}+xy_{23}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{21}^{\Hermaphrodite,s}-xy_{12}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \rho(xy_{24}^{\Hermaphrodite,s}-xy_{42}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho)xy_{23}^{\Hermaphrodite,s}+(R+\rho-r)xy_{32}^{\Hermaphrodite,s}\\
&+(R+r-\rho) xy_{14}^{\Hermaphrodite,s}+(R+\rho-r)xy_{41}^{\Hermaphrodite,s}\big{]}(1-\alpha^\Hermaphrodite)/2
\end{split}
\\

@@

\begin{split}
x_{3}^{{\Hermaphrodite}'}=&xx_{33}^{\Hermaphrodite,s}+xx_{13}^{\Hermaphrodite,s}/2+(xx_{23}^{\Hermaphrodite,s}+xx_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{23}^{\Hermaphrodite,s}-xx_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{33}^{\Hermaphrodite,s}+xy_{31}^{\Hermaphrodite,s})/2+(xy_{32}^{\Hermaphrodite,s}+xy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{34}^{\Hermaphrodite,s}-xy_{43}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite - \rho(xy_{31}^{\Hermaphrodite,s}-xy_{13}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho)xy_{32}^{\Hermaphrodite,s} +(R+\rho-r)xy_{23}^{\Hermaphrodite,s}\\
&+(R+r-\rho) xy_{41}^{\Hermaphrodite,s}+(R+\rho-r)xy_{14}^{\Hermaphrodite,s}\big{]} \alpha^\Hermaphrodite/2
\end{split}
\\

@@

\begin{split}
x_{4}^{{\Hermaphrodite}'}=&xx_{44}^{\Hermaphrodite,s}+xx_{34}^{\Hermaphrodite,s}/2+(xx_{14}^{\Hermaphrodite,s}+xx_{24}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{14}^{\Hermaphrodite,s}-xx_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{44}^{\Hermaphrodite,s}+xy_{42}^{\Hermaphrodite,s})/2+(xy_{41}^{\Hermaphrodite,s}+xy_{43}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{43}^{\Hermaphrodite,s}-xy_{34}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \rho(xy_{42}^{\Hermaphrodite,s}-xy_{24}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho)xy_{41}^{\Hermaphrodite,s}+(R+\rho-r)xy_{14}^{\Hermaphrodite,s}\\
&+(R+r-\rho) xy_{32}^{\Hermaphrodite,s}+(R+\rho-r)xy_{23}^{\Hermaphrodite,s}\big{]}(1-\alpha^\Hermaphrodite)/2
\end{split}
\\
\begin{split}
y_{1}^{{\Hermaphrodite}'}=&yy_{11}^{\Hermaphrodite,s}+yy_{13}^{\Hermaphrodite,s}/2+(yy_{12}^{\Hermaphrodite,s}+yy_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{14}^{\Hermaphrodite,s}-yy_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{11}^{\Hermaphrodite,s}+xy_{31}^{\Hermaphrodite,s})/2+(xy_{21}^{\Hermaphrodite,s}+xy_{41}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{21}^{\Hermaphrodite,s}-xy_{12}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \rho(xy_{31}^{\Hermaphrodite,s}-xy_{13}^{\Hermaphrodite,s})/2\\
&+\big{[} - (R+r+\rho)xy_{41}^{\Hermaphrodite,s} + (R+\rho-r)xy_{14}^{\Hermaphrodite,s}\\
& + (R+r-\rho)xy_{32}^{\Hermaphrodite,s} + (R+\rho-r) xy_{23}^{\Hermaphrodite,s}\big{]}\alpha^\Hermaphrodite/2
\end{split}
\\

@@

\begin{split}
y_{2}^{{\Hermaphrodite}'}=&yy_{22}^{\Hermaphrodite,s}+yy_{24}^{\Hermaphrodite,s}/2+(yy_{12}^{\Hermaphrodite,s}+yy_{23}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{23}^{\Hermaphrodite,s}-yy_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{22}^{\Hermaphrodite,s}+xy_{42}^{\Hermaphrodite,s})/2+(xy_{12}^{\Hermaphrodite,s}+xy_{32}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{12}^{\Hermaphrodite,s}-xy_{21}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \rho(xy_{42}^{\Hermaphrodite,s}-xy_{24}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho)xy_{32}^{\Hermaphrodite,s} +(R+\rho-r) xy_{23}^{\Hermaphrodite,s}\\
&+(R+r-\rho)xy_{41}^{\Hermaphrodite,s}+(R+\rho-r)xy_{14}^{\Hermaphrodite,s}\big{]}(1-\alpha^\Hermaphrodite)/2
\end{split}
\\

@@

\begin{split}
y_{3}^{{\Hermaphrodite}'}=&yy_{33}^{\Hermaphrodite,s}+yy_{13}^{\Hermaphrodite,s}/2+(yy_{23}^{\Hermaphrodite,s}+yy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{23}^{\Hermaphrodite,s}-yy_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{33}^{\Hermaphrodite,s}+xy_{13}^{\Hermaphrodite,s})/2+(xy_{23}^{\Hermaphrodite,s}+xy_{43}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{43}^{\Hermaphrodite,s}-xy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \rho(xy_{13}^{\Hermaphrodite,s}-xy_{31}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho)xy_{23}^{\Hermaphrodite,s} +(R+\rho-r)xy_{32}^{\Hermaphrodite,s}\\
&+(R+r-\rho) xy_{14}^{\Hermaphrodite,s} + (R+\rho-r) xy_{41}^{\Hermaphrodite,s}\big{]}\alpha^\Hermaphrodite/2
\end{split}
\\

@@

\begin{split}
y_{4}^{{\Hermaphrodite}'}=&yy_{44}^{\Hermaphrodite,s}+yy_{34}^{\Hermaphrodite,s}/2+(yy_{14}^{\Hermaphrodite,s}+yy_{24}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{14}^{\Hermaphrodite,s}-yy_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{44}^{\Hermaphrodite,s}+xy_{24}^{\Hermaphrodite,s})/2+(xy_{14}^{\Hermaphrodite,s}+xy_{34}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{34}^{\Hermaphrodite,s}-xy_{43}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \rho(xy_{24}^{\Hermaphrodite,s}-xy_{42}^{\Hermaphrodite,s})/2\\
&+\big{[}-(R+r+\rho) xy_{14}^{\Hermaphrodite,s} + (R+\rho-r)xy_{41}^{\Hermaphrodite,s}\\
&+(R+r-\rho) xy_{23}^{\Hermaphrodite,s} + (R+\rho-r) xy_{32}^{\Hermaphrodite,s}\big{]}(1-\alpha^\Hermaphrodite)/2.
\end{split}
\end{align}
\label{eq:recursions}
\end{subequations}

\endgroup

\noindent
The full system is therefore described by 16 recurrence equations (three diallelic loci in two sexes, $2^3 \times 2 = 16$). 
However, not all diploid types are produced under certain sex-determining systems. 
For example, with the $M$ allele fixed and an ancestral XY sex-determining system, there are XX females and XY males ($x_{3}^\Hermaphrodite=x_{4}^\Hermaphrodite=y_{3}^\male=y_{4}^\male=y_{i}^\female=0,\;\forall\; i$). 
%($xx_{11}^{\male}=xx_{12}^{\male}=xx_{22}^\male=xy_{11}^{\female}=xy_{12}^{\female}=xy_{21}^{\female}=xy_{22}^\female=yy_{11}^{\female}=yy_{12}^{\female}=yy_{22}^\female=0$). 
In this case, the system only involves six recursion equations, %because there is only one $\mathbf{M}$ locus allele and no Y-bearing female gametes. 
which we assume below to calculate the equilibria.

\newpage

\setcounter{equation}{0}
\renewcommand{\theequation}{S2.\arabic{equation}}

\section*{S2 Appendix}

\subsection*{Resident equilibria and stability}

In the resident population (allele $M$ fixed), we follow the frequency of $A$ in X-bearing female gametes, $p^\female_X$, and X-bearing male gametes, $p^\male_X$, and Y-bearing male gametes, $p^\male_Y$.
We also track the total frequency of Y among male gametes, $q$, which may deviate from $1/2$ due to meiotic drive in males. 
These four variables determine the frequencies of the six resident gamete types, which sum to one in each sex: $x_{1}^{\female}=\hat{p}_X^\female$, $x_{2}^{\female}=1-\hat{p}_X^\female$, $x_{1}^{\male}=(1-q)\hat{p}_X^\male$, $x_{2}^{\male}=(1-q)(1-\hat{p}_X^\male)$, $y_{1}^{\male}=q \hat{p}_Y^\male$, and $y_{2}^{\male}=q(1-\hat{p}_Y^\male)$. 
Mean fitnesses in the resident population are given in \nameref{tab:meanfitnesses}.

Various forms of selection can maintain a polymorphism at the $\mathbf{A}$ locus, including sexually antagonistic selection, overdominance, conflicts between diploid selection and selection upon haploid genotypes (ploidally-antagonsitic selection,~\cite{Immler:2012tl}), or a combination of these selective regimes (see below).

In particular special cases, e.g., no sex-differences in selection or meiotic drive ($s^\male=s^\female$, $h^\male=h^\female$, and $\alpha^\male=\alpha^\female=1/2$), the equilibrium allele frequency and stability can be calculated analytically without assuming anything about the relative strengths of selection and recombination. 
However, here, we focus on two regimes (tight linkage between $\mathbf{A}$ and $\mathbf{X}$ and weak selection) in order to make fewer assumptions about fitnesses.

\subsubsection*{Tight linkage between $\mathbf{X}$ and $\mathbf{A}$ loci}

We first calculate the equilibria in the ancestral population when the recombination rate between the $\mathbf{X}$ and $\mathbf{A}$ loci is small ($r$ of order $\epsilon$). 
Selection at the $\mathbf{A}$ locus will not affect evolution at the novel sex-determining locus, $\mathbf{M}$, if one allele is fixed on all backgrounds. 
We therefore focus on the five equilibria that maintain both $A$ and $a$ alleles, four of which are given to leading order by:

>>>>

%reference equations in appendices
\usepackage{xr}
\externaldocument{S1Appendix}
\externaldocument{S2Appendix}
\externaldocument{S3Appendix}

@@

%\newpage

@@

%\setcounter{equation}{0}
%\renewcommand{\theequation}{S1.\arabic{equation}}

@@

%\section*{S1 Appendix}

@@

%\subsection*{Recursion equations}

@@

@@

%In each generation we census the genotype frequencies in male and female gametes/gametophytes (hereafter, gametes) between meiosis (and any meiotic drive) and gametic competition. 
%At this stage we denote the frequencies of X- and Y-bearing gametes from males and females $x_{i}^{\Hermaphrodite}$ and $y_{i}^{\Hermaphrodite}$.
%The superscript $\Hermaphrodite \in \{\male,\female\}$ specifies the sex of the diploid that the gamete came from. 
%The subscript $i\in\{1,2,3,4\}$ specifies the genotype at the selected locus $\mathbf{A}$ and at the novel sex-determining locus $\mathbf{M}$, where $1=AM$, $2=aM$, $3=Am$, and $4=am$. 
%The gamete frequencies from each sex sum to one, $\sum_{i}x_{i}^{\Hermaphrodite}+y_{i}^{\Hermaphrodite}=1$. 

@@

%Competition then occurs among gametes of the same sex (e.g., among eggs and among sperm separately) according to the genotype at the $\mathbf{A}$ locus ($w_{1}^\Hermaphrodite=w_{3}^\Hermaphrodite=w_{A}^\Hermaphrodite$, $w_{2}^\Hermaphrodite=w_{4}^\Hermaphrodite=w_{a}^\Hermaphrodite$, see Table \ref{tab:fitnesstable}).
%The genotype frequencies after gametic competition are $x_{i}^{\Hermaphrodite,s}= w_{i}^{\Hermaphrodite}x_{i}^{\Hermaphrodite}/\bar{w}_{H}^{\Hermaphrodite}$ and $y_{i}^{\Hermaphrodite,s}= w_{i}^{\Hermaphrodite}y_{i}^{\Hermaphrodite}/\bar{w}_{H}^{\Hermaphrodite}$, where $\bar{w}_{H}^{\Hermaphrodite}=\sum_{i} w_{i}^{\Hermaphrodite}x_{i}^{\Hermaphrodite}+w_{i}^{\Hermaphrodite}y_{i}^{\Hermaphrodite}$ is the mean fitness of male ($\Hermaphrodite=\male$) or female ($\Hermaphrodite=\female$) gametes. 

@@

%Random mating then occurs between gametes to produce diploid zygotes.
%%To shorten notation we now use index $i$ (and $j$) to denote the alleles at both the $\mathbf{A}$ and $\mathbf{M}$ loci and label $MA=1$, $Ma=2$, $mA=3$, and $ma=4$, such that $i,j\in\{1,2,3,4\}$.
%The frequencies of XX zygotes are then denoted as $xx_{ij}$, XY zygotes as $xy_{ij}$, and YY zygotes as $yy_{ij}$, where $\mathbf{A}$ and $\mathbf{M}$ locus genotypes are given by $i,j\in\{1,2,3,4\}$, as above. 
%In XY zygotes, the haplotype inherited from an X-bearing gamete is given by $i$ and the haplotype from a Y-bearing gamete is given by $j$. 
%In XX and YY zygotes, individuals with diploid genotype $ij$ are equivalent to those with diploid genotype $ji$; for simplicity, we use $xx_{ij}$ and $yy_{ij}$ with $i\neq j$ to denote the average of these frequencies, $xx_{ij}=(x_{i}^{\female,s}x_{j}^{\male,s}+x_{j}^{\female,s}x_{i}^{\male,s})/2$ and $yy_{ij}=(y_{i}^{\female,s}y_{j}^{\male,s}+y_{j}^{\female,s}y_{i}^{\male,s})/2$. 
%Denoting the $\mathbf{M}$ locus genotype by $b \in \{MM, Mm, mm\}$ and the $\mathbf{X}$ locus genotype by $c \in \{XX, XY, YY\}$, zygotes develop as females with probability $k_{bc}$. 
%Therefore, the frequencies of XX females are given by $xx_{ij}^{\female}=k_{bc}xx_{ij}$, XY females are given by $xy_{ij}^{\female}=k_{bc}xy_{ij}$, and YY females are given by $yy_{ij}^{\female}=k_{bc}yy_{ij}$. 
%Similarly, XX male frequencies are $xx_{ij}^{\male}=(1-k_{bc})xx_{ij}$, XY male frequencies are $xy_{ij}^{\male}=(1-k_{bc})xy_{ij}$, and YY males frequencies are $yy_{ij}^{\male}=(1-k_{bc})yy_{ij}$.
%This notation allows both the ancestral and novel sex-determining regions to determine zygotic sex according to an XY system, a ZW system, or an environmental sex-determining system. 
%In addition, we can consider any epistatic dominance relationship between the two sex-determining loci. 
%Here, we assume that the ancestral sex-determining system ($\mathbf{X}$ locus) is XY ($k_{MMXX}=1$ and $k_{MMXY}=k_{MMYY}=0$) or ZW ($k_{MMZZ}=0$ and $k_{MMZW}=k_{MMWW}=1$) and epistatically recessive to a dominant novel sex-determining locus, $\mathbf{M}$ ($k_{Mmc}=k_{mmc}=k$). 
%Selection among diploids then occurs according to the diploid genotype at the $\mathbf{A}$ locus, $l \in \{AA, Aa, aa\}$, for an individual of type $ij$ (see Table \ref{tab:fitnesstable}). 
%The diploid frequencies after selection in sex $\Hermaphrodite$ are given by $xx_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} xx_{ij}/\bar{w}^{\Hermaphrodite}_{D}$, $xy_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} xy_{ij}/\bar{w}^{\Hermaphrodite}_{D}$, and $yy_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} yy_{ij}/\bar{w}^{\Hermaphrodite}_{D}$, where $\bar{w}^{\Hermaphrodite}_{D}= \sum_{i=1}^{4}\sum_{j=1}^{4}w_{l}^{\Hermaphrodite}xx_{ij}+w_{l}^{\Hermaphrodite}xy_{ij}+w_{l}^{\Hermaphrodite}yy_{ij}$ is the mean fitness of diploids of sex $\Hermaphrodite$. 
%Finally, these diploids undergo meiosis to produce the next generation of gametes. 
%Recombination and sex-specific meiotic drive occur during meiosis.
%%We can also assume that recombination is sex specific and/or affected by the M locus - but generally we don't so I just describe $R=R_{f}=R, r_{MM,d}=r_{Mm,d}=r_{mm,d}=r, \chi_{m}=\chi_{f}=\chi$.
%Here, we allow any relative locations for the $\mathbf{X}$, $\mathbf{A}$, and $\mathbf{M}$ loci by using three parameters to describe the recombination rates between them. 
%$R$ is the recombination rate between the $\mathbf{A}$ and $\mathbf{M}$ loci, $\rho$ is the recombination rate between the $\mathbf{M}$ and $\mathbf{X}$ loci, and $r$ is the recombination rate between the $\mathbf{A}$ and $\mathbf{X}$ loci (Fig \ref{fig:model_outline}). 
%\nameref{tab:chisubstitutions} shows replacements that can be made for each possible ordering of the loci assuming that there is no cross-over interference.
%During meiosis in sex $\Hermaphrodite$, meiotic drive occurs such that, in $Aa$ heterozygotes, a fraction $\alpha^{\Hermaphrodite}$ of gametes produced carry the $A$ allele and $(1-\alpha^\Hermaphrodite)$ carry the $a$ allele. 
%Among gametes from sex $\Hermaphrodite$, the frequencies of haplotypes (before gametic competition) in the next generation are given by
%\begingroup
%\allowdisplaybreaks
%\begin{subequations}
%\begin{align}
%%
%\begin{split}
%x_{1}^{{\Hermaphrodite}'}=&xx_{11}^{\Hermaphrodite,s}+xx_{13}^{\Hermaphrodite,s}/2+(xx_{12}^{\Hermaphrodite,s}+xx_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&-R(xx_{14}^{\Hermaphrodite,s}-xx_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
%&+(xy_{11}^{\Hermaphrodite,s}+xy_{13}^{\Hermaphrodite,s})/2+(xy_{12}^{\Hermaphrodite,s}+xy_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&- r(xy_{12}^{\Hermaphrodite,s}-xy_{21}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \rho(xy_{13}^{\Hermaphrodite,s}-xy_{31}^{\Hermaphrodite,s})/2\\
%&+\big{[}-(R+r+\rho)xy_{14}^{\Hermaphrodite,s} +(R+\rho-r)xy_{41}^{\Hermaphrodite,s}\\
%&+(R+r-\rho) xy_{23}^{\Hermaphrodite,s}+(R+\rho-r)xy_{32}^{\Hermaphrodite,s}\big{]}\alpha^\Hermaphrodite/2
%\end{split}
%\\
%%
%\begin{split}
%x_{2}^{{\Hermaphrodite}'}=&xx_{22}^{\Hermaphrodite,s}+xx_{24}^{\Hermaphrodite,s}/2+(xx_{12}^{\Hermaphrodite,s}+xx_{23}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&-R(xx_{23}^{\Hermaphrodite,s}-xx_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
%&(xy_{22}^{\Hermaphrodite,s}+xy_{24}^{\Hermaphrodite,s})/2+(xy_{21}^{\Hermaphrodite,s}+xy_{23}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
%&- r(xy_{21}^{\Hermaphrodite,s}-xy_{12}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \rho(xy_{24}^{\Hermaphrodite,s}-xy_{42}^{\Hermaphrodite,s})/2\\
%&+\big{[}-(R+r+\rho)xy_{23}^{\Hermaphrodite,s}+(R+\rho-r)xy_{32}^{\Hermaphrodite,s}\\
%&+(R+r-\rho) xy_{14}^{\Hermaphrodite,s}+(R+\rho-r)xy_{41}^{\Hermaphrodite,s}\big{]}(1-\alpha^\Hermaphrodite)/2
%\end{split}
%\\
%%
%\begin{split}
%x_{3}^{{\Hermaphrodite}'}=&xx_{33}^{\Hermaphrodite,s}+xx_{13}^{\Hermaphrodite,s}/2+(xx_{23}^{\Hermaphrodite,s}+xx_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&-R(xx_{23}^{\Hermaphrodite,s}-xx_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
%&(xy_{33}^{\Hermaphrodite,s}+xy_{31}^{\Hermaphrodite,s})/2+(xy_{32}^{\Hermaphrodite,s}+xy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&- r(xy_{34}^{\Hermaphrodite,s}-xy_{43}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite - \rho(xy_{31}^{\Hermaphrodite,s}-xy_{13}^{\Hermaphrodite,s})/2\\
%&+\big{[}-(R+r+\rho)xy_{32}^{\Hermaphrodite,s} +(R+\rho-r)xy_{23}^{\Hermaphrodite,s}\\
%&+(R+r-\rho) xy_{41}^{\Hermaphrodite,s}+(R+\rho-r)xy_{14}^{\Hermaphrodite,s}\big{]} \alpha^\Hermaphrodite/2
%\end{split}
%\\
%%
%\begin{split}
%x_{4}^{{\Hermaphrodite}'}=&xx_{44}^{\Hermaphrodite,s}+xx_{34}^{\Hermaphrodite,s}/2+(xx_{14}^{\Hermaphrodite,s}+xx_{24}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&-R(xx_{14}^{\Hermaphrodite,s}-xx_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
%&(xy_{44}^{\Hermaphrodite,s}+xy_{42}^{\Hermaphrodite,s})/2+(xy_{41}^{\Hermaphrodite,s}+xy_{43}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
%&- r(xy_{43}^{\Hermaphrodite,s}-xy_{34}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \rho(xy_{42}^{\Hermaphrodite,s}-xy_{24}^{\Hermaphrodite,s})/2\\
%&+\big{[}-(R+r+\rho)xy_{41}^{\Hermaphrodite,s}+(R+\rho-r)xy_{14}^{\Hermaphrodite,s}\\
%&+(R+r-\rho) xy_{32}^{\Hermaphrodite,s}+(R+\rho-r)xy_{23}^{\Hermaphrodite,s}\big{]}(1-\alpha^\Hermaphrodite)/2
%\end{split}
%\\
%\begin{split}
%y_{1}^{{\Hermaphrodite}'}=&yy_{11}^{\Hermaphrodite,s}+yy_{13}^{\Hermaphrodite,s}/2+(yy_{12}^{\Hermaphrodite,s}+yy_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
%&-R(yy_{14}^{\Hermaphrodite,s}-yy_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
